---
title: "NJ Q2 Distributed Solar - Interactive Dashboard"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
library(lubridate)
library(sf)
library(leaflet)
library(leafpop)
library(htmltools)
library(scales)
library(highcharter)
library(reactable)
```

```{r include = FALSE}
setwd("C:/Users/zachg/DG+Design Dropbox/4-Data and Reports/DG+ Zach WD")
```

```{r include = FALSE}
nj11 <- read_csv('NJ/NJ11.csv') 
nj11 <- nj11 %>%  rename('id' = `TI Application Number`, 'program' = 'Program', 'capacity' = 'Calculated Total System Size',
                       'segment' ='Customer Type', 'third' = 'Third Party Ownership',
                       'contractor' = 'Contractor Company', 'utility' = 'Electric Utility Name',
                       'zip' = `Premise                         Zip`, 'county' = `County                      Code`, 'type' = 
                         'Interconnection', 'date' ='Acceptance Date') %>% 
  select(id, utility, capacity, date, zip, county, segment, contractor, third, program, type)

nj22 <- read_csv('NJ/NJ22.csv')
nj22 <- nj22 %>%  rename('id' = `TI Application Number`, 'program' = 'Program', 'capacity' = 'Calculated Total System Size',
                       'segment' ='Customer Type', 'third' = 'Third Party Ownership',
                       'contractor' = 'Contractor Company', 'utility' = 'Electric Utility Name',
                       'zip' = `Premise                         Zip`, 'county' = `County                      Code`, 'type' = 
                         'Interconnection', 'date' = `Acceptance                Date`) %>% 
  select(id, utility, capacity, date, zip, county, segment, contractor, third, program, type)

nj33 <- read_csv('NJ/NJ33.csv')
nj33 <- nj33 %>% rename('id' = `SRP Registration Number`, 'program' = 'Program', 'capacity' = 'Calculated Total System Size',
                        'segment' ='Customer Type', 'third' = 'Third Party Ownership',
                        'contractor' = 'Contractor Company', 'utility' = 'Electric Utility Name',
                        'zip' = `Premise        Zip`, 'county' = `County        Code`, 'type' = 
                          'Interconnection Type', 'date' = `Registration Acceptance Date`) %>% 
  select(id, utility, capacity, date, zip, county, segment, contractor, third, program, type)

nj44 <- read_csv('NJ/NJ44.csv')
nj44 <- nj44 %>%  rename('id' = `TI Application Number`, 'program' = 'Program', 'capacity' = 'Calculated Total System Size',
       'segment' ='Customer Type', 'third' = 'Third Party Ownership',
       'contractor' = 'Contractor Company', 'utility' = 'Electric Utility Name',
       'zip' = `Premise        Zip`, 'county' = `County        Code`, 'type' = 
         'Interconnection Type', 'date' ='Acceptance Date') %>% 
  select(id, utility, capacity, date, zip, county, segment, contractor, third, program, type)

nj_2 <- rbind(nj11, nj22, nj33, nj44)
nj_2 <- distinct(nj_2)

rm(nj1, nj11, nj2, nj22, nj3, nj33, nj4, nj44)

nj_filter <- nj_2 %>%
  mutate(date = dmy(date), capacity = as.numeric(capacity)) %>% 
 # filter(date >= '2020-07-01') %>% 
  filter(type != 'Grid Supply')

nj_filter <- nj_filter %>% 
#nj_new <- nj_new %>% 
  mutate(utility = case_when(
    str_detect(utility, regex("pse", ignore_case = TRUE)) ~ "PSEG",
    str_detect(utility, regex("ac electric", ignore_case = T)) ~ "AC Electric",
    str_detect(utility, regex("jcp", ignore_case = T)) ~ "JCP&L",
    str_detect(utility, regex("rockland", ignore_case = T)) ~ 'O&R', 
    TRUE ~ 'Other (muni/co-op)'),
    quarter = zoo::as.yearqtr(date, format = "%Y-%m-%d"),
         segment_x = case_when(
           type == 'Community Solar' ~ 'Community Solar',
           segment == 'Residential' ~ 'Residential',
           TRUE ~ 'Non-Residential'
         ))
```

```{r include = FALSE}
nj_sum <- nj_filter %>% 
  filter(date >= '2020-07-01') %>% 
  group_by(segment_x, quarter) %>% 
  summarise(cap = sum(capacity)/1000)

nj_2 <- nj_filter %>% 
  filter(quarter == '2021 Q2') %>% 
  mutate(res_cap = if_else(segment_x == 'Residential', capacity, 0),
         com_cap = if_else(segment_x == 'Non-Residential', capacity, 0),
         cs_cap = if_else(segment_x == 'Community Solar', capacity, 0),
         month = months(as.Date(date)))
```

## New Jersey added **`r round(sum(nj_2$capacity)/1000,2)` MW** of new distributed solar capacity to the interconnection queue in Q2 2021

```{r echo = FALSE}
nj_sum %>%
  hchart('column', hcaes(x = as.factor(quarter), y = round(cap,2), group = segment_x)) %>% 
  hc_colors(c('#F37325', '#2490BD', "#47B970", '#1A2B40', '#b7b7b7', '#800000','#F8AA1A')) %>% 
  hc_plotOptions(series = list(stacking = 'normal')) %>% 
  hc_xAxis(title = '', labels = list(step = 1) #, minorTickInterval
  ) %>% 
  hc_yAxis(title = list(text = 'Quarterly capacity (MW)'), style = list(fontSize = "5.0vh"))
       
```

```{r include = FALSE}
nj_2_seg2 <- nj_2 %>% 
  filter(segment_x == 'Non-Residential') %>% 
  mutate(segment = as.factor(case_when(
    grepl('School', segment) ~ 'School',
    TRUE ~ segment
  ))) %>% 
  group_by(segment) %>% 
  summarise(cap = sum(capacity),
            n = n(),
            mean = mean(capacity)) %>% 
  ungroup() %>% 
  mutate(share = cap/sum(cap))
```

## Of the **`r round(sum(nj_2$capacity[nj_2$segment_x == 'Non-Residential']/1000),2)` MW** of new capacity that was Non-Residential...

```{r results = 'asis', echo = FALSE}
for (i in levels(nj_2_seg2$segment)) {
  print(paste("- ", "**", round(nj_2_seg2$cap[nj_2_seg2$segment == i]/1000,2), "**", " MW was ", i, sep = ""))
}
```

```{r echo = FALSE}
nj_2_seg2 %>% 
  hchart('pie', hcaes(x = segment, y = round(cap/1000,2)), name = "Capacity (MW)") %>% 
  hc_colors(c( '#2490BD', '#800000', '#F37325', "#47B970", '#1A2B40', '#b7b7b7', '#F8AA1A')) 

```

## Which utilities added the most capacity in Q2 2021?

```{r include = FALSE}
nj_2_ut <- nj_2 %>%
  group_by(utility) %>% 
  summarise(cap = sum(capacity))

nj_2_ut2 <- nj_2 %>%
  group_by(utility, month) %>% 
  summarise(cap = sum(capacity))
```

```{r echo = FALSE}
nj_2_ut %>% 
  hchart('pie', hcaes(x = utility, y = round(cap/1000,2)), name = "Capacity (MW)") %>% 
  hc_colors(c('#F37325', '#2490BD', "#47B970", '#b7b7b7', '#1A2B40', '#800000','#F8AA1A')) 

```

```{r echo = FALSE}
nj_2_ut2 %>% 
  mutate(month = factor(month, levels = c('April', 'May', 'June'), labels = c('April', 'May', 'June'))) %>% 
  hchart('column', hcaes(x = month, y = round(cap/1000, 2), group = utility)) %>% 
  hc_colors(c('#F37325', '#2490BD', "#47B970", '#b7b7b7', '#1A2B40', '#800000','#F8AA1A')) %>% 
  hc_plotOptions(series = list(stacking = 'normal')) %>% 
  hc_xAxis(title = '', labels = list(step = 1) #, minorTickInterval
  ) %>% 
  hc_yAxis(title = list(text = 'Monthly capacity (MW)'), style = list(fontSize = "5.0vh"))
       

```

## Which developers added the most capacity in Q2 2021?

```{r include = FALSE}
nj_2 <- nj_2 %>% 
  mutate(contractor = as.character(contractor)) %>% 
  mutate(dev_agg = case_when(
    grepl('tesla|solarcity|solar city', contractor, ignore.case = T) ~ 'Tesla/SolarCity',
    grepl('vivint|sunrun|sun run', contractor, ignore.case = T) ~ 'Vivint/SunRun',
    grepl('momentum', contractor, ignore.case = T) ~ 'Momentum Solar',
    grepl('vision', contractor, ignore.case = T) ~ 'Vision Solar',
    grepl('greenpower|green power', contractor, ignore.case = T) ~ 'Green Power Energy',
    grepl('orbit', contractor, ignore.case = T) ~ 'Orbit Energy & Power',
    grepl('sunpower|sun power', contractor, ignore.case = T) ~ 'SunPower Corporation',
    grepl('sunnymac', contractor, ignore.case = T) ~ 'SunnyMac',
    grepl('whitman', contractor, ignore.case = T) ~ 'Steve Whitman',
    grepl('evergreen', contractor, ignore.case = T) ~ 'Evergreen Energy',
    grepl('pfister', contractor, ignore.case = T) ~ 'Pfister Energy',
    grepl('national', contractor, ignore.case = T) ~ 'National Energy Partners',
    grepl('grid builder', contractor, ignore.case = T) ~ 'Grid Builders (owned by NextGrid)',
    grepl('suntuity', contractor, ignore.case = T) ~ 'Suntuity Solar',
    grepl('posigen', contractor, ignore.case = T) ~ 'Posigen NJ',
    grepl('freedom', contractor, ignore.case = T) ~ 'Freedom Forever NJ',
    grepl('entersolar', contractor, ignore.case = T) ~ 'EnterSolar',
    grepl('edf', contractor, ignore.case = T) ~ 'EDF Renewables Distributed Solutions',
    grepl('brightcore', contractor, ignore.case = T) ~ 'Brightcore',
    grepl('cents', contractor, ignore.case = T) ~ 'Solar Cents',
    TRUE ~ contractor
  ))

nj_2_dev <- nj_2 %>% 
  group_by(dev_agg, segment_x) %>% 
  summarise(cap = sum(capacity),
            n = n(),
            mean = mean(capacity)) %>% 
  ungroup() %>% 
  mutate(share = cap/sum(cap)) %>% 
  arrange(desc(cap))
```

### Top 10 developers in the residential segment:

```{r echo = FALSE}
nj_2_dev %>% filter(segment_x == 'Residential') %>%
  head(10) %>%
  select(dev_agg, cap, n, share) %>% 
  mutate(share = paste(round(share*100,2), "%", sep = ""),
         cap = format(round(as.numeric(cap), 2), big.mark = ",")) %>% 
  knitr::kable(col.names = c("Developer", "Capacity (kW)", "Number of Projects", "Market Share"),
               digits = 2, align = 'l') %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  kableExtra::add_header_above(c("Q2 2021" = 4))
```


### Top 10 developers in the commercial segment:

```{r echo = FALSE}
nj_2_dev %>% filter(segment_x == 'Non-Residential') %>%
  head(10) %>%
  select(dev_agg, cap, n, share) %>% 
  mutate(share = paste(round(share*100,2), "%", sep = ""),
         cap = format(round(as.numeric(cap), 2), big.mark = ",")) %>% 
  knitr::kable(col.names = c("Developer", "Capacity (kW)", "Number of Projects", "Market Share"),
               digits = 2, align = 'l') %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  kableExtra::add_header_above(c("Q2 2021" = 4))
```

### Top 10 developers active in both segments:

```{r include = FALSE}
nj_2_devb <- nj_2 %>% 
  group_by(dev_agg) %>%
  summarise(cap = sum(capacity),
            res = sum(res_cap),
            com = sum(com_cap),
            n = n(),
            mean = mean(capacity)) %>% 
  ungroup() %>% 
  mutate(share = cap/sum(cap)) %>% 
  filter(res != 0 & com != 0) %>% 
  arrange(desc(cap)) %>% 
  head(10)
```

```{r echo = FALSE}
nj_2_devb %>% 
  select(dev_agg, cap, res, com, mean) %>% 
  mutate(cap = format(round(as.numeric(cap), 2), big.mark = ","),
         res = format(round(as.numeric(res), 2), big.mark = ","),
         com = format(round(as.numeric(com), 2), big.mark = ","),
         mean = round(mean, 2)) %>% 
  knitr::kable(col.names = c("Developer", "Total Capacity (kW)", "Residential Capacity (kW)", "Non-Residential Capacity (kW)", "Average Total Capacity"),
               digits = 2, align = 'l') %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  kableExtra::add_header_above(c("Q2 2021" = 5))
```

## Where was new capacity added in Q2 2021?

```{r include = FALSE}
nj_map <- nj_2 %>% 
  mutate(geoid2 = as.numeric(substr(zip,1,5))) %>% 
  filter(!geoid2 == 16807|!geoid2 == 17047) %>% 
  group_by(geoid2) %>% 
  summarise(cap = sum(capacity)/1000)

njmap <- st_read('nj_shapefile/acs2019_5yr_B01003_86000US08403.shp') %>% 
  mutate(geoid2 = as.numeric(substr(geoid,8,12))) %>% 
  filter(!name == 'New Jersey',
         !geoid2 == 10990, !geoid2 == 10969)

nj__map <- left_join(njmap, nj_map, by = 'geoid2')
```

```{r include = FALSE}
zip <- read.csv('us_zip/uszips.csv')

zip_join <- zip %>% 
  select(zip, city) %>% 
  rename('geoid2' = 'zip')

nj_int <- left_join(nj__map, zip_join, by = 'geoid2') %>% 
  mutate(cap = cap*1000)

nj_int <- filter(nj_int, name != 'New Jersey')
nj_int <- mutate(nj_int, cap = round(cap,0), cap2 = if_else(is.na(cap),0,cap))

blue_mon <- c('#e0f1f9', '#2490BD') 

pal <- colorNumeric(palette = blue_mon, domain = nj_int$cap, na.color = '#FFFFFF')
labels <- sprintf("<strong>%s</strong><br/> Zip code: %s<br/> Capacity Q2 2021: %s kW",
                  nj_int$city, paste(0, nj_int$geoid2, sep = ""), comma(nj_int$cap2)) %>% 
  lapply(HTML)

map_int <- leaflet(nj_int) %>% addTiles()
```

```{r echo = FALSE}
map_int %>% addPolygons(fillColor = ~pal(cap),
                        weight = 0.5,
                        opacity = 1,
                        color = "gray",
                        dashArray = "3",
                        fillOpacity = 0.7,
                        highlightOptions = highlightOptions(color = '#2a2a2a',
                                                            weight = 2,
                                                            dashArray = "",
                                                            fillOpacity = 0.7,
                                                            bringToFront = T),
                        label = labels,
                        labelOptions = labelOptions(
                          style = list("font-weight" = "normal", padding = "3px 8px"),
                          textsize = "15px",
                          direction = "auto",
                          opacity = 0.75)) %>% 
  addProviderTiles('Esri.WorldGrayCanvas')
```



## Project highlight: Rutgers University


## Filter and download data

```{r include = FALSE}
nj_2_table <- nj_2 %>% 
  mutate(contractor = as.character(contractor)) %>% 
  mutate(Developer = case_when(
    grepl('tesla|solarcity|solar city', contractor, ignore.case = T) ~ 'Tesla/SolarCity',
    grepl('vivint|sunrun|sun run', contractor, ignore.case = T) ~ 'Vivint/SunRun',
    grepl('momentum', contractor, ignore.case = T) ~ 'Momentum Solar',
    grepl('vision', contractor, ignore.case = T) ~ 'Vision Solar',
    grepl('greenpower|green power', contractor, ignore.case = T) ~ 'Green Power Energy',
    grepl('orbit', contractor, ignore.case = T) ~ 'Orbit Energy & Power',
    grepl('sunpower|sun power', contractor, ignore.case = T) ~ 'SunPower Corporation',
    grepl('sunnymac', contractor, ignore.case = T) ~ 'SunnyMac',
    grepl('whitman', contractor, ignore.case = T) ~ 'Steve Whitman',
    grepl('evergreen', contractor, ignore.case = T) ~ 'Evergreen Energy',
    grepl('pfister', contractor, ignore.case = T) ~ 'Pfister Energy',
    grepl('national', contractor, ignore.case = T) ~ 'National Energy Partners',
    grepl('grid builder', contractor, ignore.case = T) ~ 'Grid Builders (owned by NextGrid)',
    grepl('suntuity', contractor, ignore.case = T) ~ 'Suntuity Solar',
    grepl('posigen', contractor, ignore.case = T) ~ 'Posigen NJ',
    grepl('freedom', contractor, ignore.case = T) ~ 'Freedom Forever NJ',
    grepl('entersolar', contractor, ignore.case = T) ~ 'EnterSolar',
    grepl('edf', contractor, ignore.case = T) ~ 'EDF Renewables Distributed Solutions',
    grepl('brightcore', contractor, ignore.case = T) ~ 'Brightcore',
    grepl('cents', contractor, ignore.case = T) ~ 'Solar Cents',
    TRUE ~ contractor
  )) %>% 
  select(Developer, capacity, date, utility, segment, zip) %>% 
 # mutate('Number of Projects' = 1) %>% 
  rename('Capacity' = capacity, 'Date' = date, 'Utility' = utility, 'Segment' = segment, 'Zip Code' = zip)

```


```{r echo = FALSE}

reactable(nj_2_table, filterable = TRUE,
          columns = list(Capacity = colDef(filterable = FALSE)),
          searchable = TRUE,
          showPageSizeOptions = TRUE,
          pageSizeOptions = c(5,10,20,50),
          defaultPageSize = 10)

```




